Coupled response of stomatal and mesophyll conductance to light enhances photosynthesis of shade leaves under sunflecks

### 2016-08-30 gm_pce/review_response.Rmd

---
title: "Response to referees"
author: "Court Campany"
date: "26 August 2016"
output: word_document
---

```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

*Editor*
- 1. These leaves studied in this experiment juvenille leaves, as the *E. tereticornis* trees were grown for ~ 1 year. The fact that LMA was similar within the canopy may be common in juvenille leaves and similar LMA within canopies has also been noted in Eucalpt plantations (Nouvellon et al. 2010, cited in discussion). However, we have expanded the discussion to address fixed vs variable contols of leaf physiolgy specific to LMA with additional citiations (see reply to Referee # 2 specific comments).

\
- 2. We feel that colors are necessary in this document to visualize leaf types (+light), while also showing temperature treatments.  It is very messy to distinguish all of these with symbols alone, and complicates visualization of confidence intervals on scatter plots.  We have attempted to tone-done the color palette, while still distguishing colors properly for color-blind readers.

\
- 3/4. After careful thought we have decided to move the information regarding parameters derived AC~i~ curves to the supplementary information. We only want to show that sun and shade leaves were fundamentaly different in their photosyntheitic capacity. As such, we have decided present Amax values, from these CO~2~ response curves, between the leaf types. This very simply shows the point we aim to make. If the derivation and representation of J~max~/V~cmax~ values has distracted the editor from this point, it is likely to also distract readers who are plant physiolgists. J~max~ and V~cmax~ values are still described and reported in the supplementary as additional evidence to support differences in photosynthetic capacity between leaf types.

@@

*Remko*The fitaci package used here correctly estimates J~max~.

We also acknowledge that AC~i~ response curves were intitally conceived with the notion that g~m~ was small enough to be ignored,     which we now know is untrue.  We spent consideral time attempting to fit AC~C~ curves, which generated unrealistic V~cmax~ estimates of up to ~300 mumols m^-2^s^-1^. This represents a new issue for the plant physiolgy community as more g~m~ data are generated, and we hope our datasets will eventually help contribute to a solution.

@@

\
- 2. Leaves used for gas exchange were sampled and then measured for midday water potentails. This is explained in the first sentence of the subsection "Leaf Nitrogen and hydraulic conductance" in the methods.  The method used here for measureing leaf hydraulic conductance is also one of the three methods described in Flexas et al. 2013, which compared the coordination of leaf hydraulic conductance and mesophyll conductance.

\
- 3. Due to Referee #1 concern over cuvette measured transpiration, water potentials and subsequent leaf K calculations, we re-examined the calculation and analyses in the manuscript.  Initially, leaf K was calculated with transpiration rates in shade leaves under low light conditions. As light was then increased in shade leaves and then transpiration re-measured, it is entirely possible that this change could have altered the water status of the leaf and thus subsequent water potential measurements (which we can not determine).  As such we feel that the only leaf K which can properly reported is that of sun leaves vs shade leaves under high light. The data and results in the manuscript now reflect this change. This new data allows us to address aspects of Flexas et al. 2013 (and others) mentioned above, but partially inhibits our ability to test optimization theories from Peltoniemi et al. 2012.

*Referee #2*

- 1. Wingate et al. 2007: accountd for by evan and cammereer 2012? (griffiths paper)

@@

\
- line 358:We corrected the text to match stats from Table 2

\
- line 392: We agree that thermal acclimation is not previosuly discussed before this point is made. However, the experimental design does include a +3C warming treatment, and the potential affects of warming on leaf physiology are mentioned in the introduction. As there were very few effects of climate warming observed in this aspect of the experiment, we feel it is important to address a possible explaination for the lack of observed warming effects using published findings from this experiment.

\
- Q: Nitrogen concentration in the leaves had a weak postive relationship with measured rates of mesophyll conductance if all leaves were considered. This weak relationship disappeared if sun and shade leaves were analyzed seperately. Since we can only speculate about the effect of aquaporins we have not added gm v. N in the manuscript.

\
- line 404-405: added Brooks et al. 1996 reference explaining patterns of morphology and physiology of when sun leaves become shaded.

### 2016-08-30 gm_pce/manuscript_gmes.Rmd
fixed versus variable gm

##Photosynthetic capacity and leaf photosynthesis rates 
Mean photosynthetic capacity (*A*~max~) of sun leaves  was signifcantly higher (+27%, P = 0.001) compared to shade leaves, when measured at saturating light and 25 &deg;C. Within leaf types, no effect of the warming treatment was detected on *A*~max~. *A*~max~ was positively related to leaf *N*~a~ across leaf types and temperature treatments (P = 0.01, Figure `r amaxfig`). Similarily, the the photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Supplementary Table 1).

@@

##Resource distribution and photosynthetic capacity
The allocation of *N*~a~ along canopy gradients constrains *A*~n~ and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain. Decreasing light availability should decrease the investment of N into photosynthetic enzymes within a canopy [@mooney1979environmental]. As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters *V*~cmax~ and *J*~max~ [@farquhar1980biochemical]. Our data agree with these conventional conclusions as the distribution of *N*~a~, both measures of photosynthetic capacity, and *A*~n~ were all reduced in shade leaves. Leaf mass per area (LMA), however, was not different between sun and shade leaves, consistent with small differences reported in a clonal *Eucalyptus* plantation [@nouvellon2010]. Similar LMA in sun and shade leaves could be the result of leaf formation under comparable light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types that masked differences in leaf thickness or density.

@@

\
We found that *A*~n~ and *g*~m~ scaled positively across leaf types (Fig. 3b) and, unexpectedly, increased in shade leaves when light intensity was increased to values similar in sun leaves. Possible mechanisms for the rapid (~25 min) response of *g*~m~ to light intensity include regulation of aquaporins, which can facilitate increases in the CO~2~ permeability of the cell membranes [@hanba2004overexpression; @heinen2009role; @li2014plant] or movement of chloroplasts to facilitate CO~2~ diffusion [@tholen2008]. Our findings support growing evidence that *g*~m~ is highly variable and can respond to environmental variables [@flexas2007rapid; @flexas2008mesophyll]. Here we provide empirical data showing *g*~m~ not only varies within a canopy, but that the up-regulation of *g*~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.

>>>>

##Photosynthetic capacity and leaf photosynthesis rates 
Mean photosynthetic capacity (*A*~max~) of sun leaves  was significantly higher (+27%, P = 0.001) compared to shade leaves, when measured at saturating light and 25 &deg;C. Within leaf types, no effect of the warming treatment was detected on *A*~max~. *A*~max~ was positively related to leaf *N*~a~ across leaf types and temperature treatments (P = 0.01, Figure `r amaxfig`). Similarly, the the photosynthetic parameters *J*~max~ and *V*~cmax~ were higher in sun compared to shade leaves (Supplementary Table 1).

@@

##Resource distribution and photosynthetic capacity
The allocation of *N*~a~ along canopy gradients constrains *A*~n~ and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain. Decreasing light availability should decrease the investment of N into photosynthetic enzymes within a canopy [@mooney1979environmental]. As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters *V*~cmax~ and *J*~max~ [@farquhar1980biochemical]. Our data agree with these conventional conclusions as the distribution of *N*~a~, both measures of photosynthetic capacity, and *A*~n~ were all reduced in shade leaves. Leaf mass per area (LMA), however, was not different between sun and shade leaves, consistent with small differences reported in a clonal *Eucalyptus* plantation [@nouvellon2010]. Similar LMA in sun and shade leaves could be the result of leaf formation under comparable light conditions or possible unmeasured differences in total non-structural carbohydrates contents between leaf types that masked differences in leaf thickness or density. For example, sun leaves of *Abies amabilis* foliage that became shaded exhibited physiology similar to shaded leaves, while retaining the morphology of sun leaves [@brooks1996effects].

@@

\
**We found that *A*~n~ and *g*~m~ scaled positively across leaf types (Fig. 3b) and, unexpectedly, increased in shade leaves when light intensity was increased to values similar in sun leaves. Possible mechanisms for the rapid (~25 min) response of *g*~m~ to light intensity include regulation of aquaporins, which can facilitate increases in the CO~2~ permeability of the cell membranes [@hanba2004overexpression; @heinen2009role; @li2014plant] or movement of chloroplasts to facilitate CO~2~ diffusion [@tholen2008]. Our findings support growing evidence that *g*~m~ is highly variable and can respond to environmental variables [@flexas2007rapid; @flexas2008mesophyll]. Here we provide empirical data showing *g*~m~ not only varies within a canopy, but that the up-regulation of *g*~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.**

*fixed vs variable controls of g~m~*
We found that *A*~n~ and *g*~m~ scaled positively across leaf types (Fig. 3b) and, unexpectedly, increased in shade leaves when light intensity was increased to values similar in sun leaves. Comparable LMA within the canopy likely allowed for this up-regulation of *g*~m~ , which played a critical role in the photosynthetic response of shade leaves to sunflecks. Components of LMA, such as palisade thickness and cell surface exposed to intercellular air space, set an upper limit for *g*~m~, but considerable variation in *g*~m~ can exist for a given LMA [@flexas2008mesophyll; @milla2016leaf]. Possible mechanisms for the rapid (~25 min) up-regulation of *g*~m~ to light intensity include regulation of aquaporins, which can facilitate increases in the CO~2~ permeability of the cell membranes [@hanba2004overexpression; @heinen2009role; @li2014plant] or movement of chloroplasts to facilitate CO~2~ diffusion [@tholen2008]. It has also been proposed that *g*~m~ and *K*~L~ should be coordinated, as optimizing hydraulic supply allows both *g*~s~ and *g*~m~ to enhance *A*~n~ [@flexas2013leaf]. Here, higher *K*~l~ in shade leaves under high light lends supports to the potential coordination *K*~l~ and *g*~m~ and also to the association of higher *A*~n~ with higher *K*~l~ when light becomes saturating reported by @franks2006higher.
